```{r include=FALSE}
library(knitr)
opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE,fig.width=8, fig.height=5)
```

---
title: "Using the OFB Data Extract - 1"
author: "Peter Rabinovitch"
date: "March 22, 2015"
output: html_document
---


Intro
=====
The purpose of this document is to detail some initial explorations of the OFB dataset.

Data
====

Data provided by Gary (OFB) to Trent (D4G) early March, 2015. Data consisted of two zip files, each being an extract of the sql database.  Here we work with an extract of the file DataExport.txt. For details about what was done and why, see the report 'Preparing the OFB Data Extract - 1' and the output file 


```{r}
opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE,fig.width=8, fig.height=5)
library(ggplot2)
library(lubridate)
library(tidyr)

```

We load the data, look at the fields, and the first few rows....
```{r}

data <- read.csv("./Data/OFBData.csv", stringsAsFactors=FALSE)
glimpse(data)

```


Fields
======
For each field, we display the possible values along with the count of how many rows had that value.

```{r}

data %>%
  select(Category) %>%
  group_by(Category) %>%
  tally()

data %>%
  select(CategoryLabel) %>%
  group_by(CategoryLabel) %>%
  tally()
```

Note Category and CategoryLabel are synonyms:   
A = Purchased / Donated  
B = Baby Cupboard   
C = Donated  
D = School Program  
E = City Harvest  
F = Fresh Harvest  
G = ''  

* Category A - Items that Food Bank guarantees to Agencies. Will go out and buy if none on hand. However, max quantity that is provided is determined by the designation/quota given to the specific Agency.  
* Category B - Baby Cupboard  
* Category C - Donated items: If the Food Bank has it, they will give it to the Agency. These items will not be specifically bought to match a request.  
* Category D - School Program (27 schools, Tues/Thurs)  
* Category E - City Harvest  
* Category F - Fresh Harvest Website  
* Category G - Christmas Cheer  

```{r}
data %>%
  select(Food) %>%
  group_by(Food) %>%
  tally()

data %>%
  select(Group) %>%
  group_by(Group) %>%
  tally()

```

Food is the specific item of food ordered or delivered. These have been condensed from  several fields in the original data extract, as many were eauivalent to eachother. 

Group is a grouping of food items. For example 'Fruit', 'fresh Fruit', 'Oranges and Apples', etc all get aggregated into 'Fruit'

```{r}

data %>%
  select(OrderType) %>%
  group_by(OrderType) %>%
  tally()
```

Do not know what OrderType means yet.

```{r}

data %>%
  select(CLASS) %>%
  group_by(CLASS) %>%
  tally()

data %>%
  select(Class.Description) %>%
  group_by(Class.Description) %>%
  tally()
```

Clearly CH is City Harvest, FH is Fresh Harvest, and Inv is Inventory. Need to find out what they mean, if they matter to our project.

```{r}
data %>%
  select(prodtype) %>%
  group_by(prodtype) %>%
  tally()
```
Do not yet know what prodtype means. 

Customer.Code should join to the agencies file that already has the lat & lon in it. All the Customer.Code in this file have a match in the agencies file. Note however that here are agencies that have no records in this file. For reference, the ones that do not are: A0016, A0019, A0022, A0026, A0036, A0038, A0060, A0086, A0090, A0091, A0093, A0122, A0170, A0224, T0004, T0008, T0011, T0014, T0018, T0020, X0008, X0009. The corresponding CustomerNames match too.

```{r}

data %>%
  select(Customer.Code) %>%
  group_by(Customer.Code) %>%
  tally()

data %>%
  select(CustomerClass) %>%
  group_by(CustomerClass) %>%
  tally()
```
CustomerClass may make an interesting grouping variable for the analysis.

For the numerical variables, we extract the min, max and mean for each. Deciphering the names of the columns is best achieved by looking at the code.
```{r}
data %>%
  group_by(CategoryLabel) %>%
  summarize(
  minDe=min(DeliveredQTY),
  maxDe=max(DeliveredQTY),
  meanDe=mean(DeliveredQTY),
  minO=min(Q_Ordered),
  maxO=max(Q_Ordered),
  meanO=mean(Q_Ordered),  
  minDo=min(Q_Donated),
  maxDo=max(Q_Donated),
  meanDo=mean(Q_Donated),  
  minPu=min(Q_Purchased),
  maxPu=max(Q_Purchased),
  meanPu=mean(Q_Purchased)
  ) %>% kable()
```


We'll need to find out why there is a negative value for Purchased/Donated. Again we'll need to determine the units here (cases?).

One would _assume_ that Q_Purchased + Q_Donated  = DeliveredQTY. The following plot will have a 45 degree line if it is in fact the case

```{r}
data %>%
  ggplot(aes(DeliveredQTY,Q_Purchased+Q_Donated)) +
  geom_point(size=3.2) 
```  

and we see that it is frequently the case, but not always. This will need clarification. I suspect that the points not on the 45 degree line are those where there was some donation of food that was not requested, and so it was delivered even if not ordered while the supply lasted....except for the negative value!  
  
Do not know what AMT means, I suspect it is the difference between what was ordered and delivered?
```{r}
data %>%
  group_by(CategoryLabel) %>%
  summarize(
    minAMT=min(AMT),
    maxAMT=max(AMT),
    meanAMT=mean(AMT)
  ) %>% kable()
```

Next, we have the DeliveredValue section, with DeliveredValuePerLB (really? pounds in 2015?) and the DeliveredValue. One would assume that these are in dollar values. This needs to be confirmed, as well as how the value is determined. And if they are in inflation adjusted dollars (expect not, so $1 in 1996 = (say) $0.73 in 2015).
```{r}
data %>%
  group_by(CategoryLabel) %>%
  summarize(
    minVP=min(DeliveredValuePerLB),
    maxVP=max(DeliveredValuePerLB),
    meanVP=mean(DeliveredValuePerLB),
    minV=min(DeliveredValue),
    maxV=max(DeliveredValue),
    meanV=mean(DeliveredValue)
  ) %>% kable()


```

The invoice date is next
```{r}
c(min(data$Invoice.Date),max(data$Invoice.Date))
```

```{r}
data %>%
  group_by(year(Invoice.Date),CategoryLabel) %>%
  tally()
```  

Finally, we have left in the original id variable form the source data to allow backtracking if needed.

Some Preliminary Plots
======================

In tis section we present several plots for illustrative/educational purposes. Interpretation is up to the reader.

```{r}
data %>%
  group_by(yr=year(Invoice.Date),CategoryLabel) %>%  
  summarise(
      total = sum(Q_Ordered)
  ) %>%
  ggplot(aes(x=yr,y=total,fill=CategoryLabel))+
  geom_bar(stat='identity') +
  xlab('Year') +
  ylab('Q-Ordered') +
  scale_x_continuous(labels=abbreviate,limits=c(2003,2016))
```  
  

```{r}
library(GGally)

data %>% 
  select(DeliveredQTY, Q_Ordered,Q_Donated,Q_Purchased,AMT) %>%
  sample_frac(0.1, replace=T) %>%
  ggpairs(axisLabels="none") 
  

  
data %>%
  filter(Customer.Code=='A0055', Food=="Peanut_Butter") %>%
  group_by(year(Invoice.Date),month(Invoice.Date)) %>%
  summarize(
  yr=first(year(Invoice.Date)),
  mo=first(month(Invoice.Date)),
  demand=sum(Q_Ordered)
  ) %>%
  ggplot(aes(x=(yr-2005)*12+mo-1,y=demand))+
  geom_point()  +
  xlab('Month') +
  ylab('Cases Ordered')+
  ggtitle('Agency A0055 Peanut Butter Orders')

  
  
data %>%
  filter(Q_Ordered>0, DeliveredQTY>0) %>%
  ggplot(aes(x=Q_Ordered, y=DeliveredQTY))+
  geom_point()   
  
data %>%
  filter(DeliveredQTY>0, Q_Ordered>0) %>%
  ggplot(aes(x=DeliveredQTY,y=Q_Ordered))+
  geom_bin2d(binwidth=c(1,1)) +
  xlim(0,50) +
  ylim(0,50)

```

